The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems

### 2011-10-19 shortRange.inhomo/writeDown/main.tex
revise again due to Luigi, backup

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$.  Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogenety of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
The inhomogeneous systems, in which the density of certain type of
atom changes within the length scale of several molecules, 
and thus lack of uniformity, 
the direct molecular simulation 
by the cut-off method will encounter non-trivial
difficulties.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
depend on the energy/force continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systematically
study the precision of the short-range interaction calculation, and to
develop methods for screening out the aforementioned cut-off artifact in
the inhomogeneous systems.

@@

The widely accepted definition of the force error is the \emph{root
  mean square} (RMS) \emph{error}, which is the square root of the second
moment of the error force $\mathcal E(\v r) =
\sqrt{\langle\vert\Delta\v F(\v r)\vert^2\rangle}$. This is 
calculated as:
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  =\,&
  \big\langle\sum_{j,k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \big\langle\sum_j\vert\v f^c(\v r - \v r_j)\vert^2\big\rangle +
  \big\langle\sum_{j\neq k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r'
  + \\\label{eqn:tmp4}
  \,&
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\rho(\v r', \v r'')\,\d d\v r'\d d\v r''.
\end{align}
Where  $\rho(\v r', \v r'')$ is the pair density. By using the identity
\begin{align}  \label{eqn:tmp7}
  \rho(\v r', \v r'')
  = \rho(\v r')\rho(\v r'') + C (\v r', \v r''),
\end{align}
where 
\begin{align}
C (\v r', \v r'') = [\,\rho(\v r', \v r'') -  \rho(\v r')\rho(\v r'')\,]
\end{align}
denotes the density correlation between position $\v r'$ and $\v
r''$. Eqn. \eqref{eqn:tmp4} becomes
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  = &\,
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r' \,+ \\\nonumber
  &\,
  \bigg[\int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'\,\bigg]^2 + \\\nonumber
  &\,
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\,C(\v r', \v r'')\,\d d\v r'\d d\v r'' \\\label{eqn:tmp5}
  = &\,
  \mathcal E^2_{\textrm{homo}}(\v r) +
%  \langle\Delta\v F(\v r)\,\rangle^2 +
  \mathcal E^2_{\textrm{inhomo}}(\v r) +
  \mathcal E^2_{\textrm{correlation}}(\v r).
\end{align}
If the system is homogeneous and the particle-particle correlation is
neglected, then only the first term is left on the right hand side of \eqref{eqn:tmp5}, so it
is called the homogeneity error.  This term originates from the
fluctuation of the error force $\Delta\v F(\v r)$.  The second term is
called the inhomogeneity error, because it stems from the density
inhomogeneity of the system. It is worth pointing out that the
inhomogeneity error is nothing but the magnitude of the mean error force,
namely $\mathcal E_{\textrm{inhomo}}(\v r) = \vert\langle\Delta\v F(\v
r)\,\rangle\vert$.  The last term on right hand side of \eqref{eqn:tmp5} originates from the
correlation between the particles, so it is called the correlation
error.
% The correlation error involves a double integral, which
% implies that calculating this term requires more compuational cost
% than the homogeneity and inhomogeneity errors.
The rest of this paper
assumes the 
correlation error does not dominate in the system, and can be safely
neglected.

@@

Similarly, for Eqn. \eqref{eqn:tmp3}, the Fourier mode of the the mean error
force is
\begin{align}\label{eqn:tmp17}
  \langle\Delta\v F\rangle^\wedge(\v k) =
  \hat{\v f}^c(\v k)\,
  \hat\rho(\v k),
\end{align}
where the Fourier transform of the complementary force is given by
\begin{align}\label{eqn:tmp18}
  \hat{\v f}^c(\v k) 
  & = 
  2\pi\frac{\v k}k\, i\,
  \bigg\{
  r_c^2\,u(r_c)
  \bigg[\,
  \frac{2\cos(2\pi kr)}{2\pi kr}
  -\frac{2\sin(2\pi kr)}{(2\pi kr)^2}
  \,\bigg] 
  -
  \int_{r_c}^\infty 2r\,u(r)\sin(2\pi kr)\,\d d\v r
  \bigg\}.
\end{align}

@@

In an inhomogeneous system, the inhomogeneity error may play a
dominate role in the simulation, (see examples in section~\ref{sec:tmp2}).
Since the mean error force can be calculated
by Eqn.~\eqref{eqn:tmp17} on the fly, a simple way to screen out the
inhomogeneity error is to correct the computed force by the mean error
force, namely
\begin{align}\label{eqn:define-fcorr}
  \v F_{\textrm{corr}}(\v r) = \tilde{\v F}(\v r) + \langle\Delta\v F(\v r)\rangle.
\end{align}
In this paper, the mean error force $\langle\Delta\v F(\v r)\rangle$
is also refered as \emph{correction force} in the context of
long-range force correction. This correction works because it takes
into account the force contribution outside the cut-off radius. The mean of
the corrected error force vanishes:
\begin{align}\label{eqn:mean-fcorr}
  \langle\Delta\v F_{\textrm{corr}}(\v r)\rangle
   =
  \big\langle\,
  \v F(\v r) - \tilde{\v F}(\v r) - \langle\Delta\v F(\v r)\rangle\,
  \big\rangle = 0.
\end{align}
The mean square error is given by
\begin{align} \nonumber
  \langle\vert\Delta\v F_{\textrm{corr}}(\v r)\vert^2\rangle
  & =
  \big\langle\,
  [\,\v F(\v r) - \tilde{\v F}(\v r) - \langle\Delta\v F(\v r)\rangle\,]^2
  \big\rangle \\\nonumber
  & =
  \big\langle\,
  [\,\Delta\v F(\v r) - \langle\Delta\v F(\v r)\rangle\,]^2
  \big\rangle \\\nonumber
  & =
  \langle\vert\Delta\v F(\v r)\vert^2\rangle -
  \langle\Delta\v F(\v r)\,\rangle^2 \\
  & =
  \mathcal E^2_{\textrm{homo}}(\v r) +
  \mathcal E^2_{\textrm{correlation}}(\v r),
\end{align}
containing no inhomogeneity contribution.  It is worth noticing that the mean
square error of the corrected force is nothing but the variance of the
corrected/uncorrected error force: $\langle\vert\Delta\v
F_{\textrm{corr}}(\v r)\vert^2\rangle = \textrm{Var}[\Delta\v
F_{\textrm{corr}}(\v r)] = \textrm{Var}[\Delta\v F(\v r)]$.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid/vapor
density, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional dimensionless
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  All quantities are
written in the unitless form by adding the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literatures \cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \label{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$.

@@

The liquid and gas densities of the uniform and the adaptive cut-off
methods are almost identical at the same cut-off radius, see
Fig. \ref{fig:tmp5}. This observation is also true for the surface
tension, see Fig. \ref{fig:tmp6}. This indicates that controlling the
maximum of force error (see Tab. \ref{tab:tmp1}) and introducing
larger homogeneous force error in the bulk liquid/gas regions will not
disturb the accuracy of the equilibrium densities and surface tension.
The adaptive cut-off method roughly saves 1/3 -- 1/2 of the
computational cost (see Tab. \ref{tab:tmp1}). The efficiency benefit
of the adaptive cut-off depends on the property of the system: if the
bulk regions dominate the system, and the inhomogeneity produces large
interfacial force error, then the adaptive cut-off will save non-negligible
computational expense. On the other hand, if the interfacial region is
computationally intensive, the adaptive cut-off will not improve the
efficiency greatly.

@@

Comparing the results of the force correction method at $r_c^\ast =
3.5$ and uniform cut-off radius method at $r_c^\ast = 6.0$, the force
accuracy of the former method is a little worse than the latter (see
Tab. \ref{tab:tmp1}), but the equilibrium density accuracy is better than,
and surface tension is equally good (see
Fig. \ref{fig:tmp5} and \ref{fig:tmp6}).
% In the force correction simulation, the
% inhomogeneity error is screened out and the force error term is
% dominated by the homogeneity error. And in the uniform cut-off
% simulation the force error is dominated by the inhomogeneity error.
Since the homogeneity error dominates in the former simulation and the
inhomogeneity error dominates in the latter, the above observation
indicates different natures of error play different roles in the
equilibrium densities.  The homogeneity error is actually the variation of the
corrected error force that has a vanished mean, see
Eqn. \eqref{eqn:mean-fcorr}. This implies that the corrected error
force may point to any direction with the same probability, and may cancel
in a long time average.  The inhomogeneity error,
in contrast, stems from the non-vanishing mean value of the error
force, which is always pointing to one certain direction (the opposite
direction of the density gradient), and has low  opportunity to be
self-canceled.

@@

In this section, we test the adaptive cut-off method and the long-range force
correction method in a dynamical problem: collision of nanoscale
Lennard-Jones clusters. Two initial Lennard-Jones clusters, denoted by
$A$ and $B$, are setup in the simulation box, each of which contains
the same number of particles $N_A = N_B = 10792$, with a diameter of
$d^\ast\approx 27$. The initial velocity of the clusters are $\v u_A =
\frac12(u, 0, 0)$ and $\v u_B = -\frac12(u, 0, 0)$, see
Fig. \ref{fig:tmp7}. The impact parameter $D$ is defined by the $z$
coordinate difference between the center of mass (COM) of the
clusters. The unitless impact factor defined by
\begin{align}
  x = \frac D d
\end{align}
is used to measure how far off-center the collision happens. If $x =
0$ the collision is head-on, while $x=1$ the collision is avoided.
Ref. \cite{kalweit2006collision} investigated a board range
combination of the velocity $u$ and impact factor $x$, and identified
three major collision modes: the coalescence, the stretching separation
and the shattering. The major modes were further classified into a
series submodes. In this paper, we want to study how the collision
properties are affected by the force 
precision, therefore, we focus on one special case: $u^\ast = 2.2$ and
$x^\ast = 0.6$, where the major collision mode is the coalescence, and a
poor precision of force calculation may lead to a wrong major
collision mode.

@@

We studied the liquid-vapor equilibrium and the collision of nanoscale
Lennard-Jones cluster to demonstrate the validity of the proposed
methods.  All simulation results were compared with the uniform
cut-off simulations, in which very large cut-off radii were used to
achieve real convergence.  With the same maximum RMS force error,
the precision of physical properties of the adaptive cut-off method
were the same as the uniform cut-off method. From the efficiency
point of view,
the adaptive cut-off simulation saved 34\% -- 50\% computational cost
in liquid-vapor simulation and 18\% in the cluster collision simulation.
The advantage of the adaptive cut-off method depends on the
system: if the large bulk regions are connected by high density
gradient interfaces, the adaptive cut-off method will save a large amount of
computational cost.

In the liquid-vapor equilibrium simulation, the
force correction method improved the force accuracy by one order of
magnitude without large extra computational expense. The physical
properties converged at cut-off $r_c^\ast = 3.5$, whereas the uniform
cut-off method converged at much larger cut-off radii ($r_c^\ast = 6.0$
for surface tension and $r_c^\ast = 10.0$ for equilibrium
densities). With nearly the same RMS force error, precision of
equilibrium densities of the force correction simulation were much
better than the uniform cut-off simulation.  Since the homogeneity
error dominated the force correction and the inhomogeneity error
dominated the uniform cut-off simulation, we conclude the homogeneity
error and the inhomogeneity error play different roles on the precision
of equilibrium liquid-vapor densities.  In the collision cluster case,
the force correction methods produced high-precision trajectory at a
very small cut-off radius $r_c^\ast=2.5$, with which the uniform
cut-off method was qualitatively wrong. However, due to the
inconsistency of the cluster movement and the correction force
computation, small deviations of angular velocity and momentum were
observed.

>>>>

Fortunately,
in most homogeneous systems, a satisfactory convergence can be
achieved even with a very small cut-off value, for example,
$2.5\sigma$.  Moreover, it is possible to correct the systematic error
of the potential energy, the pressure and the free energy by applying
the standard long range correction (LRC)~\cite{allen87a}, which
assumes the uniformity outside the cut-off radius and integrates the
contribution of the mentioned properties to infinity. 
% The spacial inhomogenety of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
Instead for inhomogeneous systems, in which the density of certain type of
atom changes within the length scale of several molecules, 
and thus lack of uniformity, 
the direct molecular simulation 
by the cut-off method will encounter non-trivial
difficulties.
The simulation results are found to converge only at a very large
cut-off radius value~\cite{duque2004some, ismail2007application}, also
depend on the energy/force continuity at the cut-off and the way of
applying the long-range corrections~\cite{trokhymchuk1999computer,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte,
  shen2007comparative}.  Therefore, it is necessary to systematically
study the precision of the short-range interaction calculation, and to
develop methods for screening out the aforementioned cut-off artifact in
the inhomogeneous systems.

@@

The widely accepted definition of the force error is the \emph{root
  mean square} (RMS) \emph{error}, which is the square root of the second
moment of the error force $\mathcal E(\v r) =
\sqrt{\langle\vert\Delta\v F(\v r)\vert^2\rangle}$. This is 
calculated as:
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  =\,&
  \big\langle\sum_{j,k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \big\langle\sum_j\vert\v f^c(\v r - \v r_j)\vert^2\big\rangle +
  \big\langle\sum_{j\neq k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r'
  + \\\label{eqn:tmp4}
  \,&
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\rho(\v r', \v r'')\,\d d\v r'\d d\v r''.
\end{align}
Where  $\rho(\v r', \v r'')$ is the pair density. By using the identity
\begin{align}  \label{eqn:tmp7}
  \rho(\v r', \v r'')
  = \rho(\v r')\rho(\v r'') + C (\v r', \v r''),
\end{align}
where 
\begin{align}
C (\v r', \v r'') = [\,\rho(\v r', \v r'') -  \rho(\v r')\rho(\v r'')\,]
\end{align}
denotes the density correlation between position $\v r'$ and $\v
r''$. Eqn. \eqref{eqn:tmp4} becomes
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  = &\,
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r' \,+ \\\nonumber
  &\,
  \bigg[\int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'\,\bigg]^2 + \\\nonumber
  &\,
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\,C(\v r', \v r'')\,\d d\v r'\d d\v r'' \\\label{eqn:tmp5}
  = &\,
  \mathcal E^2_{\textrm{homo}}(\v r) +
%  \langle\Delta\v F(\v r)\,\rangle^2 +
  \mathcal E^2_{\textrm{inhomo}}(\v r) +
  \mathcal E^2_{\textrm{correlation}}(\v r).
\end{align}
If the system is homogeneous and the particle-particle correlation is
neglected, then only the first term is left on the right hand side of \eqref{eqn:tmp5}, so it
is called the homogeneity error.  This term originates from the
fluctuation of the error force $\Delta\v F(\v r)$.  The second term is
called the inhomogeneity error, because it stems from the density
inhomogeneity of the system. It is worth pointing out that the
inhomogeneity error is nothing but the magnitude of the mean error force,
namely $\mathcal E_{\textrm{inhomo}}(\v r) = \vert\langle\Delta\v F(\v
r)\,\rangle\vert$.  The last term on right hand side of \eqref{eqn:tmp5} originates from the
correlation between the particles, so it is called the correlation
error.
The correlation error involves a double integral, which
implies that the calculation  requires more compuational resources
than the homogeneity and inhomogeneity errors.
The rest of this paper
assumes the 
correlation error does not dominate in the system, and can be safely
neglected.
It is worth pointing out that this is not always true, for example, in
critical region where the corrlation plays an important role in the system,
the corrlation error should be taken into account.

@@

Similarly, for Eqn. \eqref{eqn:tmp3}, the Fourier mode of the the mean error
force is (in terms of Fourier transform):
\begin{align}\label{eqn:tmp17}
  \langle\Delta\v F\rangle^\wedge(\v k) =
  \hat{\v f}^c(\v k)\,
  \hat\rho(\v k),
\end{align}
where the Fourier transform of the complementary force is given by
\begin{align}\label{eqn:tmp18}
  \hat{\v f}^c(\v k) 
  & = 
  2\pi\frac{\v k}k\, i\,
  \bigg\{
  r_c^2\,u(r_c)
  \bigg[\,
  \frac{2\cos(2\pi kr)}{2\pi kr}
  -\frac{2\sin(2\pi kr)}{(2\pi kr)^2}
  \,\bigg] 
  -
  \int_{r_c}^\infty 2r\,u(r)\sin(2\pi kr)\,\d d\v r
  \bigg\}.
\end{align}

@@

In an inhomogeneous system, the inhomogeneity error may play a
dominate role in the simulation, (see examples in section~\ref{sec:tmp2}).
Since the mean error force can be calculated
by Eqn.~\eqref{eqn:tmp17} on the fly, a simple way to screen out the
inhomogeneity error is to correct the computed force by the mean error
force, namely
\begin{align}\label{eqn:define-fcorr}
  \v F_{\textrm{corr}}(\v r) = \tilde{\v F}(\v r) + \langle\Delta\v F(\v r)\rangle.
\end{align}
In this paper, the mean error force $\langle\Delta\v F(\v r)\rangle$
is also refered as \emph{correction force} in the context of
long-range force correction. This correction works because it takes
into account the force contribution outside the cut-off radius. The mean of
the corrected error force vanishes:
\begin{align}\label{eqn:mean-fcorr}
  \langle\Delta\v F_{\textrm{corr}}(\v r)\rangle
   =
  \big\langle\,
  \v F(\v r) - \tilde{\v F}(\v r) - \langle\Delta\v F(\v r)\rangle\,
  \big\rangle = 0.
\end{align}
The mean square error is given by
\begin{align} \nonumber
  \langle\vert\Delta\v F_{\textrm{corr}}(\v r)\vert^2\rangle
  & =
  \big\langle\,
  [\,\v F(\v r) - \tilde{\v F}(\v r) - \langle\Delta\v F(\v r)\rangle\,]^2
  \big\rangle \\\nonumber
  & =
  \big\langle\,
  [\,\Delta\v F(\v r) - \langle\Delta\v F(\v r)\rangle\,]^2
  \big\rangle \\\nonumber
  & =
  \langle\vert\Delta\v F(\v r)\vert^2\rangle -
  \langle\Delta\v F(\v r)\,\rangle^2 \\
  & =
  \mathcal E^2_{\textrm{homo}}(\v r) +
  \mathcal E^2_{\textrm{correlation}}(\v r),
\end{align}
containing no inhomogeneity contribution.  It is worth noticing that the mean
square error of the corrected force is nothing but the variance of the
error force: $\langle\vert\Delta\v
F_{\textrm{corr}}(\v r)\vert^2\rangle = \textrm{Var}[\Delta\v
F_{\textrm{corr}}(\v r)] = \textrm{Var}[\Delta\v F(\v r)]$.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid/vapor
density, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional Lennard-Jones
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  All quantities are
rescaled by the Lennard-Jones units and added the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literatures \cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \label{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$.

@@

The liquid and gas densities of the uniform and the adaptive cut-off
methods are almost identical at the same cut-off radius, see
Fig. \ref{fig:tmp5}. This observation is also true for the surface
tension, see Fig. \ref{fig:tmp6}. This indicates that controlling the
maximum of force error (see Tab. \ref{tab:tmp1}) and introducing
larger homogeneous force error in the bulk liquid/gas regions will not
disturb the accuracy of the equilibrium densities and surface tension.
The adaptive cut-off method roughly saves 1/3 -- 1/2 of the
computational cost (see Tab. \ref{tab:tmp1}). The efficiency benefit
of the adaptive cut-off depends on the property of the system: if the
bulk regions dominate the system, and the inhomogeneity produces large
interfacial force error, then the adaptive cut-off will save non-negligible
computational resources. On the other hand, if the interfacial region is
computationally intensive, the adaptive cut-off will not improve the
efficiency greatly.

@@

Comparing the results of the force correction method at $r_c^\ast =
3.5$ and uniform cut-off radius method at $r_c^\ast = 6.0$, the force
accuracy of the former method is a little worse than the latter (see
Tab. \ref{tab:tmp1}), but the equilibrium density accuracy is better than,
and surface tension is equally good (see
Fig. \ref{fig:tmp5} and \ref{fig:tmp6}).
% In the force correction simulation, the
% inhomogeneity error is screened out and the force error term is
% dominated by the homogeneity error. And in the uniform cut-off
% simulation the force error is dominated by the inhomogeneity error.
Since the homogeneity error dominates in the former simulation and the
inhomogeneity error dominates in the latter, the above observation
indicates different natures of error play different roles in the
equilibrium densities.  The homogeneity error is actually the variation of the
corrected error force with its mean equal to zero (see
Eqn.~\eqref{eqn:mean-fcorr}). This implies that the corrected error
force may point to any direction with the same probability, and may cancel
in a long time average.  The inhomogeneity error,
in contrast, stems from the non-vanishing mean value of the error
force, which is always pointing to one certain direction (the opposite
direction of the density gradient), and has low  opportunity to be
self-canceled.

@@

In this section, we test the adaptive cut-off method and the long-range force
correction method in a dynamical problem: collision of nanoscale
Lennard-Jones clusters. Two initial Lennard-Jones clusters, denoted by
$A$ and $B$, are setup in the simulation box, each of which contains
the same number of particles $N_A = N_B = 10792$, with a diameter of
$d^\ast\approx 27$. The initial velocity of the clusters are $\v u_A =
\frac12(u, 0, 0)$ and $\v u_B = -\frac12(u, 0, 0)$, see
Fig. \ref{fig:tmp7}. The impact parameter $D$ is defined by the $z$
coordinate difference between the center of mass (COM) of the
clusters. The dimensionless impact factor defined by
\begin{align}
  x = \frac D d
\end{align}
is used to measure how far off-center the collision happens. If $x =
0$ the collision is head-on, while $x=1$ the collision is avoided.
Ref. \cite{kalweit2006collision} investigated a board range
combination of the velocity $u$ and impact factor $x$, and identified
three major collision modes: the coalescence, the stretching separation
and the shattering. The major modes were further classified into a
series submodes. In this paper, we want to study how the collision
properties are affected by the force 
precision, therefore, we focus on one special case: $u^\ast = 2.2$ and
$x^\ast = 0.6$, where the major collision mode is the coalescence, and a
poor precision of force calculation may lead to a wrong major
collision mode.

@@

We studied the liquid-vapor equilibrium and the collision of nanoscale
Lennard-Jones cluster to demonstrate the validity of the proposed
methods.  All simulation results were compared with the uniform
cut-off simulations, in which very large cut-off radii were used to
achieve real convergence.  With the same maximum RMS force error,
the precision of physical properties of the adaptive cut-off method
were the same as the uniform cut-off method. From the efficiency
point of view,
the adaptive cut-off simulation saved 34\% -- 50\% computational cost
in liquid-vapor simulation and 18\% in the cluster collision simulation.
The advantage of the adaptive cut-off method depends on the
system: if the large bulk regions are connected by high density
gradient interfaces, the adaptive cut-off method will save a large amount of
computational resources.

In the liquid-vapor equilibrium simulation, the
force correction method improved the force accuracy by one order of
magnitude without large extra computational expense. The physical
properties converged at cut-off $r_c^\ast = 3.5$, whereas the uniform
cut-off method converged at much larger cut-off radii ($r_c^\ast = 6.0$
for surface tension and $r_c^\ast = 10.0$ for equilibrium
densities). With nearly the same RMS force error, precision of
equilibrium densities of the force correction simulation were much
better than the uniform cut-off simulation.  Since the homogeneity
error dominated the force correction and the inhomogeneity error
dominated the simulation with uniform cut-off, we conclude the homogeneity
error and the inhomogeneity error play different roles on the precision
of equilibrium liquid-vapor densities.  In the collision cluster case,
the force correction methods produced high-precision trajectory at a
very small cut-off radius $r_c^\ast=2.5$, with which the uniform
cut-off method was qualitatively wrong. However, due to the
inconsistency of the cluster movement and the correction force
computation, small deviations of angular velocity and momentum were
observed.
